Methods paper for erosion

### 2017-10-23 docs/methods/erosion_methods.tex
worked on numerics and formulation sections

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral Equation Formulation} 
\label{sec:bies}
Here we describe a boundary integral equation for the governing
equations.   This requires solving the incompressible Stokes equations
with Dirichlet boundary conditions 
\begin{alignat*}{3}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Omega, \\
  \uu &= \ff,  && \xx \in \bd\Omega.
\end{alignat*}
and also computing several quantities of interest, including the shear
stress.  The viscosity $\mu$ sets the time scale, and we will assume
that $\mu=1$.  We start by defining the double-layer potential, which is
the convolution of the stresslet with a density function
$\eeta:\bd\Omega \rightarrow \RR^2$,
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:velocityDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the unit outward
normal at $\yy$.  For any density function $\eeta$, the double-layer
potential satisfies the incompressible Stokes equations, but it can not
capture velocity fields corresponding to rigid body
motions~\cite{pow-mir1987}.  A standard method to remove this rank
deficiency is to complement the double-layer potential with a Stokeslets
and rotlets centered inside each interior body
\begin{align}
  S[\llambda_k,\cc_k] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_k
  \quad \text{and} \quad
  R[\xi_k,\cc_k] = \frac{\rr_k^\perp}{\rho^2}\xi_k, 
  \quad k=1,\ldots,M,
  \label{eqn:stokeslet_rotlet}
\end{align}
where  $\rr = \xx - \cc_k$, $\cc_k$ is a point inside the $k^{th}$ body,
$\rr^\perp = (r_2,-r_1)$, and $\rho = \|\rr\|$.  Then, any solution of
the Stokes equations with a Dirichlet boundary condition $\ff$ can be
written in terms of the completed double-layer potential
\begin{align}
  \label{eqn:completed_DLP}
  \uu(\xx) = \DD[\eeta](\xx) + \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
              \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:completed_BIE}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
      \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx) + \NN_0[\eeta](\xx),
      \quad \xx \in \bd\Omega, 
      \label{eqn:DLP}\\
  \llambda_k &= \frac{1}{2\pi}\int_{\gamma_k} \eeta(\yy) \, ds_\yy,
  \quad k=1,\ldots,M,
  \label{eqn:stokeslet} \\
  \xi_k &= \frac{1}{2\pi}\int_{\gamma_k} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy, \quad k=1,\ldots,M,
  \label{eqn:rotlet}
\end{align}
\end{subequations}
where 
\begin{align*}
  \NN_0[\eeta](\xx) = \int_{\Gamma} 
    (\nn(\xx) \otimes \nn(\yy))\eeta(\yy) \, ds_\yy,
\end{align*}
is being used to remove the rank one null space that results from the
compatibility constraint that results from the flux-free requirement of
$\ff$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Shear Stress}
The shear stress is defined by
\begin{align*}
  \tau = -\left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
\end{align*}
where $\ss$ and $\nn$ are the unit tangent and normal vectors,
respectively (Figure~\ref{fig:schematic}).  The velocity field of the
completed double-layer potential~\eqref{eqn:completed_DLP} involves
three terms: the double-layer potential $\DD$, the Stokeslets $S$, and
the rotlets $R$.  We first compute the deformation tensor $\ssigma =
\frac{1}{2}(\nabla\uu + \nabla\uu^T)$ for each of these terms
individually.

The deformation tensor of the double-layer
potential~\eqref{eqn:velocityDLP} at $\xx \in \Omega$ is
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) \eeta(\yy) ds_\yy.
  \end{aligned}
\end{equation}
We require the deformation tensor on $\bd\Omega$, and this requires
taking the limit of equation~\eqref{eqn:shearStressDLP} when $\xx$ tends
to $\xx_0 \in \bd\Omega$.  The limiting value is given
by~\cite{qua-bir2014a}
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0), \quad \xx_0 \in \bd\Omega,
\end{align*} 
where
\begin{align}
  J[\eeta](\xx_0) = \frac{1}{2}\pderiv{\eeta}{\ss} \cdot \ss
  \left[ \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right].
  \label{eqn:shearStressJump}
\end{align}
The deformation tensor of the Stokeslets and
rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{equation}
  \label{eqn:shearStressSR}
  \begin{aligned}
  \ssigma^S(\xx_0) &= \sum_{k=1}^{M}
    \frac{\rr \cdot \llambda_k}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\sum_{k=1}^M
    \frac{\xi_k}{\rho^4} \left(\rr \otimes \rr^\perp + 
    \rr^\perp \otimes \rr \right),
  \end{aligned}
\end{equation}
where $\rr = \xx - \cc_k$, $\cc_k$ is the point chosen inside body
$\gamma_k$, and $\rho = \|\rr_k\|$.  Now that the deformation tensor of
each component of the completed double-layer potential is computed, the
shear stress at $\xx_0 \in \bd\Omega$ is
\begin{align*}
  \tau = -2 \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss.
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Vorticity}
For the Stokes equations, on boundaries that have a no-slip boundary
condition, the shear stress is equivalent to the vorticity $w = v_x -
u_y$.  Therefore, to compute the $w(\xx)$ for $\xx \in
\overline{\Omega}$, we only need to compute $w(\xx)$ for $\xx \in
\Omega$.  As we did for the shear stress, we compute the vorticity due
to the double-layer potential, Stokeslets, and rotlets individually.
The vorticity of the double-layer potential~\eqref{eqn:velocityDLP} at
$\xx \in \Omega$ is
\begin{align}
  w^{\DD}(\xx) = -\frac{1}{\pi}\int_{\bd\Omega} 
    \frac{(\rr \cdot \nn^\perp) + (\rr \cdot \nn)}{\rho^4}
    (\rr \cdot \eeta) ds_{\yy}.
  \label{eqn:vorticityDLP}
\end{align}
The vorticity due to the Stokeslet is
\begin{align*}
  w^S(\xx) = -\frac{1}{\pi} \sum_{k=1}^{M} 
    \frac{\rr \cdot \llambda_k^\perp}{\rho^2},
\end{align*}
and the vorticity of a rotlet is zero.  Therefore, the vorticity at $\xx
\in \Omega$ is
\begin{align*}
  w(\xx) = w^\DD(\xx) + w^S(\xx).
\end{align*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Drag}
\todo[inline]{check all signs in this section.  Fortran code assumes a
normal pointing into the eroding body and a tangent vector that
corresponds to a couterclockwise parameterization}
Finally, with the pressure and shear stress, the drag on a single body
with boundary $\gamma_k$ is
\begin{align}
  F_D = \int_{\gamma_k} (-p \nn + \tau \ss)\,ds.
  \label{eqn:drag}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods\label{s:method}} 
There are three main numerical methods that need to be developed:~the
boundary integral equation solver for
equation~\eqref{eqn:completed_BIE}, computing the quantities of interest
in Section~\ref{sec:qois}, and updating the shape of the eroding bodies
with equation~\eqref{eqn:bodyEvolution}.  So that we can form stable
simulations with long time horizons, we use numerical methods that
achieve spectral accuracy in space, second-order accuracy in time, and
maintain an equispaced discretization of the bodies.

In Section~\ref{sec:spatial_discretization}, we describe how the
geometry is represented.  Then, Sections~\ref{sec:on_boundary}
and~\ref{sec:off_boundary} describe how the velocity and other
quantities of interest are computed at points on and off the boundary,
respectively.  Then, in Section~\ref{sec:thetaL}, methods used to update
the shape are described.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial Discretization}
\label{sec:spatial_discretization}
Each body is represented with at $N_\iin$ points as the Fourier series
\begin{align*}
  \xx^k(\theta) = \sum_{j = -N/2}^{N/2-1} \hat{x}_j^k e^{ij\theta},
    \quad k=1,\ldots,M.
\end{align*}
The exterior geometry is represented with a Fourier series using
$N_\out$ points.  Derivatives of the shape are computed in Fourier space
\begin{align*}
  \frac{d\xx^k(\theta)}{d\theta} = 
    \sum_{j = -N/2}^{N/2-1} ij\hat{x}_j^k e^{ij\theta},
    \quad k=1,\ldots,M.
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quadrature: On Boundary Calculations} 
\label{sec:on_boundary}
We start by discussing the quadrature methods that are used to compute
the velocity~\eqref{eqn:DLP}, shear stress~\ref{eqn:shearStressDLP},
pressure~\eqref{eqn:pressureDLP}, and drag~\eqref{eqn:drag} at target
points on the boundary of the domain.  The velocity on $\bd\Omega$ is
required to solve the integral equation~\eqref{eqn:completed_BIE} for
the density function, Stokeslets and rotlets.  Once these are computed,
computing the shear stress, pressure, and drag is a post-processing
step.

All these integrals involve integrands that are both periodic and
smooth, so the trapezoid rule guarantees spectral
accuracy~\cite{tre-wei2014}.  However, before the trapezoid rule can be
applied, singularity subtraction and odd-even
integration~\cite{sid-isr1988} must be used.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Velocity}
To numerically evaluate the completed double-layer
potential~\eqref{eqn:completed_BIE}, we define $N = MN_\iin + N_\out$ to
be the total number of discretization points of $\bd\Omega$.  The
double-layer potential in~\eqref{eqn:DLP} is discretized with the
trapezoid rule as
\begin{align}
  \DD[\eeta](\xx_i) = \sum_{j=1}^{N} K(\xx_i,\xx_j) \eeta(\xx_j) 
      \Delta s_{j}, \quad i=1,\ldots,N,
  \label{eqn:trapDLP}
\end{align}
where $\Delta s_j$ is the arclength term of $\bd\Omega$ at
$\xx_j$, and
\begin{align*}
  K(\xx,\yy) = \frac{1}{\pi} \frac{\rr \cdot \nn}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2},
\end{align*}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$, is the kernel of the
double-layer potential.  The diagonal term of~\eqref{eqn:trapDLP} is
replaced with the limiting value of the double-layer potential
\begin{align*}
  D(\xx,\xx) = -\frac{\kappa(\xx)}{2\pi}(\ss(\xx) \otimes \ss(\xx)),
\end{align*}
where $\kappa(\xx)$ is the curvature and $\ss(\xx)$ is the unit tangent
vector at $\xx \in \bd\Omega$.  The strengths of the
Stokeslets~\eqref{eqn:stokeslet} and the rotlets~\eqref{eqn:rotlet} are
also discretized with the trapezoid rule.  Applying this discretization
strategy to~\eqref{eqn:completed_BIE}, the dense linear system that
needs to be solved is
\begin{equation}
  \label{eqn:completed_BIE_discretized}
  \begin{aligned}
    \ff(\xx_i) &= -\frac{1}{2}\ssigma(\xx_i) + \sum_{j=1}^{N} 
      K(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_{j} + 
      \sum_{k=1}^M S[\llambda_k,\cc_k](\xx_i) \\
      &\hspace{40pt}
      +\sum_{k=1}^M R[\xi_k,\cc_k](\xx_i) 
      +\sum_{\xx_j \in \Gamma} (\nn(\xx_i) \otimes
      \nn(\xx_j))\eeta(\xx_j), \\
    \llambda_k &= \sum_{\xx_i \in \gamma_k} \eeta(\xx_i) 
      \Delta s_i \\ 
    \mu_k &= \sum_{\xx_i \in \gamma_k}
      {(\xx_i^k - \cc_k)}^\perp \cdot \eeta(\xx_i^k) \Delta s_i^k,
  \end{aligned}
\end{equation}
for $i=1,\ldots,N$, $k=1,\ldots,M$.  Since all the integrands are smooth
and periodic, by using the trapezoid rule, we are guaranteed that the
numerical solutions converges to the true solution with spectral
accuracy.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral Equation Formulation} 
\label{sec:bies}
Here we describe a boundary integral equation for the governing
equations.   This requires solving the incompressible Stokes equations
with Dirichlet boundary conditions 
\begin{alignat*}{3}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Omega, \\
  \uu &= \ff,  && \xx \in \bd\Omega,
\end{alignat*}
and also computing several quantities of interest, including the shear
stress.  The viscosity $\mu$ sets the time scale, and we will assume
that $\mu=1$.  We start by defining the double-layer potential, which is
the convolution of the stresslet with a density function
$\eeta:\bd\Omega \rightarrow \RR^2$,
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:velocityDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the unit outward
normal at $\yy$.  For any density function $\eeta$, the double-layer
potential satisfies the incompressible Stokes equations, but it can not
capture velocity fields corresponding to rigid body
motions~\cite{pow-mir1987}.  A standard method to remove this rank
deficiency is to complement the double-layer potential with a Stokeslets
and rotlets centered inside each interior body
\begin{align}
  S[\llambda_k,\cc_k] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_k
  \quad \text{and} \quad
  R[\xi_k,\cc_k] = \frac{\rr^\perp}{\rho^2}\xi_k, 
  \quad k=1,\ldots,M,
  \label{eqn:stokeslet_rotlet}
\end{align}
where  $\rr = \xx - \cc_k$, $\cc_k$ is a point inside the $k^{th}$ body,
$\rr^\perp = (r_2,-r_1)$, and $\rho = \|\rr\|$.  Then, any solution of
the Stokes equations with a Dirichlet boundary condition $\ff$ can be
written in terms of the completed double-layer potential
\begin{align}
  \label{eqn:completed_DLP}
  \uu(\xx) = \DD[\eeta](\xx) + \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
              \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:completed_BIE}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
      \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx) + \NN_0[\eeta](\xx),
      \quad \xx \in \bd\Omega, 
      \label{eqn:DLP}\\
  \llambda_k &= \frac{1}{2\pi}\int_{\gamma_k} \eeta(\yy) \, ds_\yy,
  \quad k=1,\ldots,M,
  \label{eqn:stokeslet} \\
  \xi_k &= \frac{1}{2\pi}\int_{\gamma_k} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy, \quad k=1,\ldots,M.
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Here,
\begin{align*}
  \NN_0[\eeta](\xx) = \int_{\Gamma} 
    (\nn(\xx) \otimes \nn(\yy))\eeta(\yy) \, ds_\yy
\end{align*}
is being used to remove the rank one null space that results from the
compatibility constraint that results from the flux-free requirement of
$\ff$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Shear Stress}
The shear stress is
\begin{align*}
  \tau = -\left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
\end{align*}
where $\ss$ and $\nn$ are the unit tangent and normal vectors,
respectively (Figure~\ref{fig:schematic}).  The velocity field of the
completed double-layer potential~\eqref{eqn:completed_DLP} involves
three terms: the double-layer potential $\DD$, the Stokeslets $S$, and
the rotlets $R$.  We first compute the deformation tensor $\ssigma =
\frac{1}{2}(\nabla\uu + \nabla\uu^T)$ for each of these terms
individually.

The deformation tensor of the double-layer
potential~\eqref{eqn:velocityDLP} at $\xx \in \Omega$ is
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) \eeta(\yy) ds_\yy.
  \end{aligned}
\end{equation}
We require the deformation tensor on $\bd\Omega$, and this requires
taking the limit of equation~\eqref{eqn:shearStressDLP} when $\xx$ tends
to $\xx_0 \in \bd\Omega$.  The limiting value is given
by~\cite{qua-bir2014a}
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0), \quad \xx_0 \in \bd\Omega,
\end{align*} 
where
\begin{align}
  J[\eeta](\xx_0) = \frac{1}{2}\left(\pderiv{\eeta}{\ss} 
    \cdot \ss\right) \left[ 
  \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right].
  \label{eqn:shearStressJump}
\end{align}
The deformation tensor of the Stokeslets and
rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{equation}
  \label{eqn:shearStressSR}
  \begin{aligned}
  \ssigma^S(\xx_0) &= \sum_{k=1}^{M}
    \frac{\rr \cdot \llambda_k}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\sum_{k=1}^M
    \frac{\xi_k}{\rho^4} \left(\rr \otimes \rr^\perp + 
    \rr^\perp \otimes \rr \right),
  \end{aligned}
\end{equation}
where $\rr = \xx - \cc_k$, $\cc_k$ is the point chosen inside body
$\gamma_k$, and $\rho = \|\rr_k\|$.  Now that the deformation tensor of
each component of the completed double-layer potential is computed, the
shear stress at $\xx_0 \in \bd\Omega$ is
\begin{align*}
  \tau = -2 \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss.
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Vorticity}
For the Stokes equations, on boundaries that have a no-slip boundary
condition, the shear stress is equivalent to the vorticity $w = v_x -
u_y$.  Therefore, to compute $w(\xx)$ for $\xx \in \overline{\Omega}$,
we only need to compute $w(\xx)$ for $\xx \in \Omega$.  As we did for
the shear stress, we compute the vorticity due to the double-layer
potential, Stokeslets, and rotlets individually.  The vorticity of the
double-layer potential~\eqref{eqn:velocityDLP} at $\xx \in \Omega$ is
\begin{align}
  w^{\DD}(\xx) = -\frac{1}{\pi}\int_{\bd\Omega} 
    \frac{(\rr \cdot \nn^\perp) + (\rr \cdot \nn)}{\rho^4}
    (\rr \cdot \eeta) ds_{\yy}.
  \label{eqn:vorticityDLP}
\end{align}
The vorticity due to the Stokeslet is
\begin{align*}
  w^S(\xx) = -\frac{1}{\pi} \sum_{k=1}^{M} 
    \frac{\rr \cdot \llambda_k^\perp}{\rho^2},
\end{align*}
and the vorticity of a rotlet is zero.  Therefore, the vorticity at $\xx
\in \Omega$ is
\begin{align*}
  w(\xx) = w^\DD(\xx) + w^S(\xx).
\end{align*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Drag}
\todo[inline]{check all signs in this section.  Fortran code assumes a
normal pointing into the eroding body and a tangent vector that
corresponds to a couterclockwise parameterization}
With the pressure and shear stress, the drag on a single body with
boundary $\gamma_k$ is
\begin{align}
  F_D = \int_{\gamma_k} (-p \nn + \tau \ss)\,ds.
  \label{eqn:drag}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods\label{s:method}} 
There are three main numerical methods that need to be developed:~the
boundary integral equation solver for
equation~\eqref{eqn:completed_BIE}, computing the quantities of interest
in Section~\ref{sec:qois}, and updating the shape of the eroding bodies
with equation~\eqref{eqn:bodyEvolution}.  So that we can form stable
simulations with long time horizons, we use numerical methods that
achieve spectral accuracy in space, fourth-order accuracy in time, and
maintain an equispaced discretization of the bodies.

In Section~\ref{sec:spatial_discretization}, we describe how the
geometry is represented.  Then, Sections~\ref{sec:on_boundary}
and~\ref{sec:off_boundary} describe how the velocity and other
quantities of interest are computed at points on and off the boundary,
respectively.  Section~\ref{sec:fmm} describes how the layer potentials
are computed efficiently.  Finally, in Section~\ref{sec:thetaL}, methods
used to update the shape are described.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial Discretization}
\label{sec:spatial_discretization}
Each body is represented with $N_\iin$ points as the Fourier series
\begin{align*}
  \xx^k(\theta) = \sum_{j = -N_\iin /2}^{N_\iin /2-1} \hat{x}_j^k e^{ij\theta},
    \quad k=1,\ldots,M.
\end{align*}
The exterior geometry is represented with a Fourier series using
$N_\out$ points.  Derivatives of the shape are computed in Fourier space
\begin{align*}
  \frac{d\xx^k(\theta)}{d\theta} = 
    \sum_{j = -N_\out /2}^{N_\out /2-1} ij\hat{x}_j^k e^{ij\theta},
    \quad k=1,\ldots,M.
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quadrature: On Boundary Calculations} 
\label{sec:on_boundary}
We start by discussing the quadrature methods that are used to compute
the velocity~\eqref{eqn:DLP}, shear stress~\eqref{eqn:shearStressDLP},
pressure~\eqref{eqn:pressureDLP}, and drag~\eqref{eqn:drag} at target
points on the boundary of the domain.  The velocity on $\bd\Omega$ is
required to solve the integral equation~\eqref{eqn:completed_BIE} for
the density function, Stokeslets and rotlets.  Once these are computed,
computing the shear stress, pressure, and drag is a post-processing
step.  All these integrals involve integrands that are periodic.  For
smooth integrands, the trapezoid rule guarantees spectral
accuracy~\cite{tre-wei2014}.  However, some of the integrands contain a
singularity whose strength will be reduced with singularity subtraction,
and then spectral accuracy is achieved with odd-even
integration~\cite{sid-isr1988} must be used.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Velocity}
To numerically evaluate the completed double-layer
potential~\eqref{eqn:completed_BIE}, we le $N = MN_\iin + N_\out$ be the
total number of discretization points of $\bd\Omega$.  The double-layer
potential in~\eqref{eqn:DLP} is discretized with the trapezoid rule as
\begin{align}
  \DD[\eeta](\xx_i) = \sum_{j=1}^{N} K(\xx_i,\xx_j) \eeta(\xx_j) 
      \Delta s_{j}, \quad i=1,\ldots,N,
  \label{eqn:trapDLP}
\end{align}
where $\Delta s_j$ is the arclength term of $\bd\Omega$ at
$\xx_j$, and
\begin{align*}
  K(\xx,\yy) = \frac{1}{\pi} \frac{\rr \cdot \nn}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2},
\end{align*}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$, is the kernel of the
double-layer potential.  The diagonal term of~\eqref{eqn:trapDLP} is
replaced with the limiting value of the double-layer potential
\begin{align*}
  D(\xx,\xx) = -\frac{\kappa(\xx)}{2\pi}(\ss(\xx) \otimes \ss(\xx)),
\end{align*}
where $\kappa(\xx)$ is the curvature and $\ss(\xx)$ is the unit tangent
vector at $\xx \in \bd\Omega$.  The strengths of the
Stokeslets~\eqref{eqn:stokeslet} and the rotlets~\eqref{eqn:rotlet} are
also discretized with the trapezoid rule.  Applying this discretization
strategy to~\eqref{eqn:completed_BIE}, the dense linear system that
needs to be solved is
\begin{equation}
  \label{eqn:completed_BIE_discretized}
  \begin{aligned}
    \ff(\xx_i) &= -\frac{1}{2}\ssigma(\xx_i) + \sum_{j=1}^{N} 
      K(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_{j} + 
      \sum_{k=1}^M S[\llambda_k,\cc_k](\xx_i) \\
      &\hspace{40pt}
      +\sum_{k=1}^M R[\xi_k,\cc_k](\xx_i) 
      +\sum_{\xx_j \in \Gamma} (\nn(\xx_i) \otimes
      \nn(\xx_j))\eeta(\xx_j), \\
    \llambda_k &= \sum_{\xx_i \in \gamma_k} \eeta(\xx_i) 
      \Delta s_i, \\ 
    \mu_k &= \sum_{\xx_i \in \gamma_k}
      {(\xx_i - \cc_k)}^\perp \cdot \eeta(\xx_i) \Delta s_i,
  \end{aligned}
\end{equation}
for $i=1,\ldots,N$, $k=1,\ldots,M$.  Since all the integrands are smooth
and periodic, by using the trapezoid rule, we are guaranteed that the
numerical solutions converges to the true solution with spectral
accuracy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Shear Stress}
We have decomposed the deformation tensor into the contributions due to
three different terms.  The contributions from the Stokeslets and the
rotlets can be directly evaluated since $\rr \neq 0$ for all $\xx \in
\gamma$.  However, the kernel of the deformation tensor due to the
double-layer potential~\eqref{eqn:shearStressDLP} 
\begin{align*}
  K_\ssigma(\xx,\yy)\eeta(\yy) &= \frac{1}{2\pi} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn)
    \right. \\ & \hspace{25pt}+ \left.
    \frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right),
\end{align*}
has a singularity that is asymptotic to $\bigO(\rho^{-2})$ when $\yy
\rightarrow \xx$.  The singularity strength can be reduced since the
velocity corresponding to a constant density function is
constant~\cite{poz1992}.  Therefore, the deformation tensor of the
double-layer potential is equivalent to
\begin{align*}
  \ssigma^{\DD}(\xx) = \int_{\bd\Omega}K_\ssigma(\xx,\yy)
      (\eeta(\yy) - \eeta(\xx)) \, ds_{\yy}, \quad \xx \in \bd\Omega.
\end{align*}
This expression for the deformation tensor has a singularity that is
asymptotic to $\bigO(\rho^{-1})$ when $\yy \rightarrow \xx$.  Finally,
to achieve spectral accuracy, we apply odd-even
integration~\cite{sid-isr1988}.  That is, the deformation tensor due to
the double-layer potential at a discretization point $\xx_i$ on body
$\gamma_k$, with $i$ being odd, is approximated as
\begin{align}
  \ssigma^{\DD}(\xx_i) = \sum_{\xx_j \notin \gamma_k}
    K_\ssigma(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_j + 
  \sum_{\substack{\xx_j \in \gamma_k \\ j \: \mathtt{even}}}
    K_\ssigma(\xx_i,\xx_j) (\eeta(\xx_j) - \eeta(\xx_i)) \Delta s_j,
  \label{eqn:stressOddEven}
\end{align}
and a similar approximation is used when $i$ is even.  The deformation
tensor due to the Stokeslets and rotlets is given in
equation~\eqref{eqn:shearStressSR}.  Finally, the jump in the
deformation tensor~\eqref{eqn:shearStressJump} is added, and the
required arclength derivative is computed in Fourier space.  Then, the
shear stress can be computed by computing the deformation tensor in the
normal direction, and taking the inner product with the tangent
direction.
